Summary: Dissertation Proposal

### 2018-11-13 r/Graphing+Analysis.Rmd
11/12

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(startR)
library(here)
library(tidyverse)
library(RColorBrewer)
library(corrplot)
library(nlme)
library(pls)
library(factoextra)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X)

# stop<-sum_id%>%filter(species_code=="stop")
# qup<-sum_id%>%filter(species_code=="qup")
# rbp<-sum_id%>%filter(species_code=="rbp")
```

@@

```{r,echo=FALSE}
sample <- group_by(sum_id, site, species, phase) %>% summarize(n=n())
barplot(table(sum_id$species, sum_id$site), main="Sample size by phase and species",
   legend = rownames(counts), beside=TRUE, las=2)
```

@@

### Distributions
```{r,warning=FALSE,echo=FALSE}
predictors <- sum_site %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity) %>% 
  gather() %>% 
  na.omit()

ggplot(predictors,aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```
Site-level predictors are not normally distributed....

### Correlations
```{r,echo=FALSE}
cor = cor(sum_id %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,consp_bm,consp_den,
         carn_bm,carn_den,cca_cover,lc_cover,
         ma_cover, ma_canopy,ta_cover,ta_canopy,rugosity), 
  use="pairwise.complete.obs")
corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)

```

@@

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~tot_bm+tot_den+scar_bm+scar_den+consp_bm+consp_den+carn_bm+carn_den+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Fish Community Variables", na.action = na.omit)
```

Variables carn_den and carn_bm are essentially identical - I will remove carn_den and use carn_bm as the main preditor metric. Total fish biomass and density (tot_bm and tot_den) are highly correlated with both scarid and carnivore variables, so I will remove them. conspific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove consp_den?). The selected variables I will proceed with are as follows:
```{r,echo=FALSE}
pairs(~scar_bm+scar_den+consp_den+carn_bm+rugosity, data=sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main = "Selected Fish Community Variables", na.action = na.omit)
```

Scarid biomass, carnivore biomass, and rugosity should be integrated via pcA. Scarid density and conspific density should also be combined, or one should be omitted.

```{r,echo=FALSE}
pairs(~cca_cover+lc_cover+ma_cover+ma_canopy+ta_cover+ta_canopy+rugosity, data = sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main ="Benthic Community Variables", na.action = na.omit)
```

All benthic variables are highly correlated. I can omit cca_cover and ma_cover (accounted for by ma_canopy). The rest should be combined via pcA. First I will assess remaining correlations between selected benthic and fish community variables:

```{r,echo=FALSE}
pairs(~scar_bm+scar_den+carn_bm+lc_cover+ta_cover+ma_canopy+ta_canopy+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

Unsure how to proceed here if I do end up using both benthic and fish community variables as predictors, because some are corellated (esp. scarid_bm.s and all benthic variables....). It looks like I would integrate scar_bm and carn_bm into a pcA with all other benthic/rugosity variables. Variables scar_den and consp_den are relatively independent of all other variables (except each other).

Current outline for site-level pcA tasks:
If benthic=predictor:
- lc_cover + ta_cover + ma_canopy + ta_canopy + rugosity + scar_bm + carn_bm
- scar_den + consp_den
If benthic=response:
- rugosity + scar_bm + carn_bm
- lc_cover + ta_cover + ma_canopy + ta_canopy
- scar_den + consp_den

@@

### Corellations
```{r,echo=FALSE}
cor=cor(sum_id %>% 
          select(g_frac,fr,br,for_dur,for_bites), use="pairwise.complete.obs")
corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)
```

```{r,echo=FALSE}
pairs(~g_frac+fr+br+for_dur+for_bites, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Grazing Response Variables",na.action = na.omit)
```

@@

@@

# PCAs -->
*Do I need to normalize my variables pre-pcA? Currently just scaled, except for grazing response variables which are log-transformed* -->

## Predictors -->

### All predictors -->
```{r}
pca <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity,scar_bm,scar_den,tot_bm)),
  center = TRUE,scale. = TRUE)
summary(pca)

plot(pca,type="l")

fviz_pca_biplot(pca,
                col.var = "black",
                col.ind = as.factor(sum_site$island), # color by groups
                palette = c("#00AFBB", "#E7B800", "#FC4E07"),
                addEllipses = T, ellipse.level=0.95,
                ellipse.type = "confidence",
                label = "var",
                legend.title = "Groups",
                repel = T
                ) +
  scale_color_brewer(palette="Dark")+
  theme_minimal()

fviz_mca_biplot(pca, 
  habillage = sum_id$island, addEllipses = TRUE,
  label = "var", shape.var = 15) +
  scale_color_brewer(palette="Dark3")+
  theme_minimal()
```

<!-- ### Benthic variables -->
<!-- ```{r, echo=FALSE} -->
<!-- pred.benthic <- sum_site %>% select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
<!-- pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)  -->

>>>>

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(startR)
library(here)
library(tidyverse)
# library(lazyeval)
# library(ellipse)
library(RColorBrewer)
#library(ggbiplot)
library(corrplot)
library(nlme)
library(pls)
library(factoextra)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor")

sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor") %>%
  select(island,site,spearfishing,diving,species,species_code,phase,length_cm,g_frac,fr,br,tdb,for_dur,for_bites,tot_bm,tot_den,scar_bm,herb_bm,scar_den,conspec_bm,conspec_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,cover_ta,canopy_ma,canopy_ta,rugosity,int_tot,int_scar,int_con,int_hetscar,int_dam)

stop<-sum_id%>%filter(species_code=="stop")
qup<-sum_id%>%filter(species_code=="qup")
rbp<-sum_id%>%filter(species_code=="rbp")
```

@@

```{r,echo=FALSE}
sample <- group_by(sum_id, site, species, phase) %>% summarize(n=n())
ggplot(data=sample, aes(x=site, y=n, fill=phase)) +
    geom_bar(stat="identity", position=position_dodge()) + facet_grid(.~species) + theme_minimal() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_fill_grey()
# counts <- table(sum_id$species, sum_id$site)
# barplot(counts, main="Sample size by phase and species",
#   legend = rownames(counts), beside=TRUE, las=2)
```

@@

### Distributions
```{r,warning=FALSE,echo=FALSE}
predictors <- sum_site %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,canopy_ma,cover_ta,canopy_ta,rugosity) %>% 
  gather() %>% 
  na.omit()

ggplot(predictors,aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

### Correlations
```{r,echo=FALSE}
site_var <- sum_id %>% select(tot_bm,tot_den,scar_bm,scar_den,conspec_bm,conspec_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,canopy_ma,cover_ta,canopy_ta,rugosity)
cor = cor(site_var, use="pairwise.complete.obs")
my_colors <- brewer.pal(5, "Spectral")
my_colors = colorRampPalette(my_colors)(100)
ord <- order(cor[1, ])
cor_ord = cor[ord, ord]
plotcorr(cor_ord , col=my_colors[cor_ord*50+50] , mar=c(1,1,1,1)  )
```

@@

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~tot_bm+tot_den+scar_bm+scar_den+conspec_bm+conspec_den+carn_bm+carn_den+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Fish Community Variables", na.action = na.omit)
```

Variables carn_den and carn_bm are essentially identical - I will remove carn_den and use carn_bm as the main preditor metric. Total fish biomass and density (tot_bm and tot_den) are highly correlated with both scarid and carnivore variables, so I will remove them. Conspecific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove conspec_den?). The selected variables I will proceed with are as follows:
```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar_bm+scar_den+conspec_den+carn_bm+rugosity, data=sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main = "Selected Fish Community Variables", na.action = na.omit)
```

Scarid biomass, carnivore biomass, and rugosity should be integrated via pcA. Scarid density and conspecific density should also be combined, or one should be omitted.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~cover_cca+cover_lc+cover_ma+canopy_ma+cover_ta+canopy_ta+rugosity, data = sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main ="Benthic Community Variables", na.action = na.omit)
```

All benthic variables are highly correlated. I can omit cover_cca and cover_ma (accounted for by canopy_ma). The rest should be combined via pcA. First I will assess remaining correlations between selected benthic and fish community variables:

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar_bm+scar_den+carn_bm+cover_lc+cover_ta+canopy_ma+canopy_ta+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

Unsure how to proceed here if I do end up using both benthic and fish community variables as predictors, because some are corellated (esp. scarid_bm.s and all benthic variables....). It looks like I would integrate scar_bm and carn_bm into a pcA with all other benthic/rugosity variables. Variables scar_den and conspec_den are relatively independent of all other variables (except each other).

Current outline for site-level pcA tasks:
If benthic=predictor:
- cover_lc + cover_ta + canopy_ma + canopy_ta + rugosity + scar_bm + carn_bm
- scar_den + conspec_den
If benthic=response:
- rugosity + scar_bm + carn_bm
- cover_lc + cover_ta + canopy_ma + canopy_ta
- scar_den + conspec_den

@@

### Corellations
```{r,echo=FALSE}
grazing <- sum_id %>% select(g_frac,fr,br,for_dur,for_bites)
cor=cor(grazing, use="pairwise.complete.obs")
ord <- order(cor[1, ])
cor_ord = cor[ord, ord]
plotcorr(cor_ord , col=my_colors[cor_ord*50+50] , mar=c(1,1,1,1)  )
```

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~g_frac+fr+br+for_dur+for_bites, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Grazing Response Variables",na.action = na.omit)
```

@@

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~g_frac+fr+br+for_dur, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Grazing Response Variables",na.action = na.omit)
```

@@

<!-- # pcAs -->
<!-- *Do I need to normalize my variables pre-pcA? Currently just scaled, except for grazing response variables which are log-transformed* -->

<!-- ## Predictors -->

<!-- ### All predictors -->
<!-- ```{r} -->
<!-- predictors <- sum_site %>% select(cover_lc,cover_ta,canopy_ma,canopy_ta,rugosity,scar_bm,scar_den,tot_bm) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
<!-- pca <- prcomp(na.omit(predictors),center = TRUE,scale. = TRUE)  -->
<!-- summary(pca) -->

<!-- plot(pca,type="l") -->
<!-- ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->
<!-- ggsave("pcApred.png") -->
<!-- ``` -->
<!-- ### Benthic variables -->
<!-- ```{r, echo=FALSE} -->
<!-- pred.benthic <- sum_site %>% select(cover_lc,cover_ta,canopy_ma,canopy_ta,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
<!-- pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)  -->

<!-- ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE, varname.size=2.5, varname.adjust=1.1) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->

### 2018-11-13 r/SummaryOrals.Rmd
11/12

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
library(car)
library(reshape2)
library(EnvStats)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("data2.csv", stringsAsFactors=FALSE) %>% filter(!(Species.Code=="qup" & Site=="Turtle Bay")) %>% filter(!(Species.Code=="qup" & Site=="Pallaster East")) # remove due to low sample size

>>>>

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
# library(ggbiplot)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
library(car)
library(reshape2)
library(EnvStats)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("data2.csv", stringsAsFactors=FALSE) %>% filter(!(Species.Code=="qup" & Site=="Turtle Bay")) %>% filter(!(Species.Code=="qup" & Site=="Pallaster East")) # remove due to low sample size
